### 2016-08-01 2015CompleteDerivation/pictures/independentDof.tex
Adding a figure

\documentclass{standalone}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{calculator}
\begin{document}
	\begin{tikzpicture}[scale=.4]

	% INPUTS

	\newcommand{\xvar}{5};					% x component of omega
	\newcommand{\yvar}{6};					% y component of omega	
	\newcommand{\xprimevar}{6.4};	    % length of the bottom side of the second triangle, called x prime
	% Note: must have  x' < omega = sqrt(x^2 + y^2)

	\newcommand{\dvar}{20}; 				% space between the two drawings

	% outer rectangle and diagram name
	\draw (-8,-12) rectangle (34,14);
	\node at (13,-8) {\huge Independent d.o.f.};

	% outer rectangle and diagram name, alternate method, needs more calculations...
	%		\draw (-\yvar - 2,-12) rectangle (\dvar + \dvar, \yvar + \xvar + 2);
	%		\node at (13,-8) {\huge Independent d.o.f.};

	% calculate omega from x and y components
	\SQUARE{\xvar}{\xsquared};
	\SQUARE{\yvar}{\ysquared};
	\ADD{\xsquared}{\ysquared}{\sum};	
	\SQUAREROOT{\sum}{\omegalength};

	% calculate the angle that omega creates w.r.t. the horizontal axis
	\DIVIDE{\yvar}{\xvar}{\sol};	
	\ARCTAN{\sol}{\newsol};
	\RADtoDEG{\newsol}{\thetavar};

	% calculate y prime and theta prime from x prime 
	\SQUARE{\omegalength}{\omegasquared};
	\SQUARE{\xprimevar}{\xprimesquared};
	\SUBTRACT{\omegasquared}{\xprimesquared}{\dif};
	\SQUAREROOT{\dif}{\yprimevar};

	\DIVIDE{\yprimevar}{\xprimevar}{\sol};	
	\ARCTAN{\sol}{\newsol};
	\RADtoDEG{\newsol}{\thetaprimevar};

	% calculations
	\newcommand{\phivar}{90-\thetavar};							  % second angle 
	\newcommand{\phiprimevar}{90-\thetaprimevar};	 	   % second prime angle

	% drawing the right basic triangle 
	\filldraw[fill=red,opacity=0.5] (0,0) -- (\xvar,0) -- (\xvar,\yvar) -- cycle;

	\filldraw[fill=red,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black] (0,0) rectangle (\xvar,-\xvar) node[pos=.5] {\large $ {dq^1}\wedge {dp_1}$};

	\filldraw[fill=red,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black] (\xvar,0) rectangle (\xvar + \yvar, \yvar) node[pos=.5] {\large ${dq^2}\wedge {dp_2}$};

	\filldraw[fill=red,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar:(0,0)}]  (0,0) rectangle (\omegalength, \omegalength) node[pos=.5, rotate=\thetavar] {\large $\omega$};

	% drawing the complex right triangle  
	\filldraw[black,fill=yellow,opacity=0.5,rotate around={\thetavar:(\dvar,0)}] (\dvar + \omegalength,0) -- (\dvar,0) -- ++(-\thetaprimevar: \xprimevar) -- ++(\phiprimevar:\yprimevar);

	\filldraw[black,fill=yellow,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar - \thetaprimevar:(\dvar,0)}]  (\dvar,0) rectangle (\dvar + \xprimevar,-\xprimevar) node[pos=.5, rotate=\thetavar-\thetaprimevar] {\large $d\hat{q}^1\wedge d\hat{p}_1$};

	\filldraw[black,fill=yellow,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar - \thetaprimevar:(\dvar,0)}]  (\dvar + \xprimevar,0) rectangle (\dvar + \xprimevar + \yprimevar, \yprimevar) node[pos=.5, rotate=\thetavar-\thetaprimevar] {\large $d\hat{q}^2\wedge d\hat{p}_2$};

	\filldraw[black,fill=yellow,fill opacity=0.1,draw opacity=.2,text opacity =1,text=black,rotate around={\thetavar:(\dvar,0)}]  (\dvar,0) rectangle (\dvar + \omegalength, \omegalength) node[pos=.5, rotate=\thetavar] {\large $\omega$};

	\end{tikzpicture}

\end{document}